Preserving Reciprocal Consistency in \\ Distributed Graph Databases

### 2020-04-20 slides/main.tex
corrections

% ----------------------------------------------------------------------------------------
%	Configuration
%----------------------------------------------------------------------------------------

@@

\usetheme[titleformat=regular,numbering=none,progressbar=frametitle]{metropolis}

% newcastle colour scheme
\definecolor{nclBlue}{HTML}{003A65}
\definecolor{nclRed}{HTML}{D91A35}

\setbeamercovered{transparent}
\setbeamercovered{invisible}
\setbeamercolor{progress bar}{fg=nclRed}
\setbeamercolor{frametitle}{bg=nclBlue}
\setbeamercolor{background canvas}{bg=white}
\setbeamercolor{block body}{bg=gray!30}
\setbeamercolor{block title}{bg=nclBlue,fg=white}
\setbeamercolor{palette primary}{fg=white, bg=nclBlue}

\linespread{1.5}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{url}
\usepackage{amsmath}
\usepackage{mathspec}
\usepackage[font=footnotesize]{caption}

% Thin fonts
\setsansfont[BoldFont={Fira Sans}, Numbers={OldStyle}]{Fira Sans Light}
\setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Light}

% Thick fonts
% \setsansfont[BoldFont={Fira Sans SemiBold}, Numbers={OldStyle}]{Fira Sans}
% \setmathsfont(Digits)[Numbers={Lining, Proportional}]{Fira Sans Book}

% tikz configuration
\usetikzlibrary{shapes.geometric, arrows}
\usetikzlibrary{arrows, positioning, shapes}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{calc}

\tikzstyle{vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
\tikzstyle{mini_vertex} = [circle, minimum width=15pt,draw,inner sep=0pt]
\tikzstyle{server} = [rectangle, rounded corners,minimum width=3cm,minimum height=4cm,draw=black]
\tikzstyle{mini_server} = [rectangle, rounded corners,minimum width=1.5cm,minimum height=1.9cm,draw=black]
\tikzstyle{local} = [thick,->,>=stealth]
\tikzstyle{dist} = [thick,dashed,->,>=stealth]
\tikzstyle{bignode}=[draw, circle, minimum width=70pt,align=center]

%----------------------------------------------------------------------------------------
%	Title + Overview Slide
% ----------------------------------------------------------------------------------------

\title[]{Preserving Reciprocal Consistency in \\ Distributed Graph Databases}

\author[Waudby \& Ezhilchelvan \& Webber \& Mitrani]{\underline{\emph{Jack Waudby}}$^{1}$, Paul Ezhilchelvan$^{1}$, Jim Webber$^{2}$ \& Isi Mitrani$^{1}$}
\institute[NCL]
{
  $^{1}$Newcastle University\\
  $^{2}$Neo4j
}
\date{April 27, 2020}

@@

\begin{frame}
\maketitle
\begin{tikzpicture}[overlay, remember picture]
  \node[above left=2.8cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_newcastle_uni.png}};
  \node[above left=1.2cm and 0.852cm of current page.south east]{\includegraphics[width=3cm]{images/logo_neo4j.png}};
\end{tikzpicture}
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
% Background
% ------------------------------------------------

@@

\begin{frame}
  \frametitle{Graph Databases}
    \begin{figure}
      \includegraphics[scale=0.5]{images/property_graph.png}
      \caption{Labeled property graph\footnotemark}
    \end{figure}
  \footnotetext[1]{
    \url{https://neo4j.com}
  }
\end{frame}

\begin{frame}
  \frametitle{Graph Databases}
  \begin{itemize}
    \uncover<1->{
    \item \textbf{Efficient graph analysis:} reachability, pattern matching, shortest path search and clustering
    }
    \uncover<2->{
    \item \textbf{Use cases:} telecommunications, pharma, publishing, finance, social media
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Graph Database Popularity}
  \begin{figure}[h!]
    \centering
    \includegraphics[scale=0.3]{images/db_popularity.png}
    \caption{Database popularity by data model\footnotemark}
  \end{figure}
  \footnotetext[1]{
    \url{https://db-engines.com/en/ranking_categories}
  }
\end{frame}

\begin{frame}
  \frametitle{Distributed Graph Databases}
  \begin{itemize}
    \uncover<1->{
    \item Graph exceeds the storage capacity of a single machine
    }
    \uncover<2->{
    \item Partition the graph across multiple machines in a cluster, with partitions replicated for fault-tolerance and availability
    }
    \uncover<3->{
    \item Graph partitioning inevitably leads to \textbf{distributed edges}
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Distributed Graph Databases: Partitioning}
\input{./figures/graph}
\end{frame}

\begin{frame}
  \frametitle{Distributed Graph Databases: Partitioning}
  \begin{figure}[h!]
    \centering
    \resizebox{0.6\textwidth}{!}{
    \begin{tikzpicture}[node distance=2cm]
      \node (s1) [server,label=$S_1$] {};
      \node (s2) [server, right of=s1,xshift=2cm,yshift=2cm,label=$S_2$] {};
      \node (s3) [server, right of=s1,xshift=6cm,label=$S_3$] {};

      \node (v1) [vertex,xshift=-0.5cm,yshift=-1.5cm] {};
      \node (v2) [vertex, above of=v1,xshift=-0.3cm,yshift=1cm] {};
      \node (v3) [vertex, above of=v1,xshift=0.6cm] {};
      \node (v4) [vertex, below right of=v3,xshift=-0.5cm,yshift=-0.3cm] {};
      \node (v5) [vertex, above right of=v3,xshift=-0.6cm,yshift=-0.5cm] {};
      \node (v6) [vertex, right of=v5,xshift=0.3cm,yshift=0.8cm] {};
      \node (v7) [vertex, below of=v6, xshift=1cm,yshift=0.4cm] {};
      \node (v8) [vertex, right of=v4, yshift=-0.3cm,xshift=4.5cm] {};
      \node (v9) [vertex, above of=v8,xshift=-0.4cm,yshift=-0.5cm] {};
      \node (v10) [vertex, above of=v9,xshift=-2.5cm,yshift=-0.3cm] {};
      \node (v11) [vertex, right of=v9,yshift=0.5cm,xshift=-0.3cm] {};
      \node (v12) [vertex, above of=v11,yshift=-1cm,xshift=-1cm] {};
      \node (v13) [vertex, above right of=v6,yshift=-0.4cm,xshift=-0.1cm] {};
      \draw [local] (v2) -- (v1);
      \draw [local] (v1) -- (v3);
      \draw [local] (v3) -- (v2);
      \draw [local] (v2) -- (v5);
      \draw [local] (v1) -- (v4);
      \draw [local] (v5) -- (v4);
      \draw [dist] (v5) -- (v6);
      \draw [dist] (v4) -- (v7);
      \draw [dist] (v8) -- (v4);
      \draw [local] (v8) -- (v9);
      \draw [local] (v9) -- (v11);
      \draw [local] (v10) -- (v7);
      \draw [dist] (v10) -- (v12);
      \draw [dist] (v12) -- (v13);
      \draw [local] (v10) -- (v6);
      \draw [dist] (v10) -- (v11);
      \draw [local] (v13) -- (v10);
      \draw [local] (v11) -- (v12);
    \end{tikzpicture}
  }
\caption{Graph partitioned across 3 servers}
  \end{figure}
\end{frame}

%------------------------------------------------
% Problem 1: Reciprocal Consistency}
% ------------------------------------------------

\begin{frame}
  \frametitle{Reciprocal Consistency}
  \begin{center}
  \textbf{A logical edge is represented by 2 physical records}
  \end{center}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}[node distance=7cm]
      \uncover<1-> {
      \node (n1) [bignode] {\texttt{Flight}};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:AVAILABLE}} (n2);
    }
      \uncover<2-> {
         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Seat Available In Flight}};
         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Flight Has Seat Available}};
         \draw [<->,dashed](q1) -- (q2);
      }
    \end{tikzpicture}
    \caption{A reciprocal consistent edge}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Reciprocal Consistency Violation}
  Two concurrent transactions:
  \begin{itemize}
  \item Mr Red requests to book the seat, $T_R$
  \item Mr Blue requests to book the seat, $T_B$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Reciprocal Consistency Violation}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}[node distance=2cm]
      \uncover<1-> {
        \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
        \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
        \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
      }
      \uncover<2-3> {
        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
      }
      \uncover<3-4> {
        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
      }
      \uncover<4-> {
        \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
      }
      \uncover<5-> {
        \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
      }
    \end{tikzpicture}
    \caption{Concurrent transactions interleave and violate reciprocal consistency}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Reciprocal Consistency Violation}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}[node distance=7cm]
      \node (n1) [bignode] {\texttt{Flight}};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
      \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
      \node (n2) [bignode, right of=n1] {\texttt{Seat}};
      \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);

      \uncover<2-> {
      \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
      \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
      \draw [<->,dashed](q1) -- (q2);
      \draw [<->,dashed](q1) -- (q2);
      \coordinate (a) at (4,-2);
      \coordinate (b) at (3,-3);
      \coordinate (c) at (3,-2);
      \coordinate (d) at (4,-3);
      \draw [very thick,red] ($ (a) $) -- ($ (b) $);
      \draw [very thick,red] ($ (c) $) -- ($ (d) $);
    }
    \end{tikzpicture}
    \caption{Reciprocal inconsistent distributed edge}
  \end{figure}
\end{frame}

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}[node distance=2cm]
%       \uncover<1->{
%         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
%         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
%         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
%         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
%         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
%       }
%       \uncover<2->{
%         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
%         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
%         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
%         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
%         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
%       }
%     \end{tikzpicture}
%     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
%   \end{figure}
% \end{frame}

\begin{frame}
  \frametitle{Distributed Edge Reciprocal Consistency}
  \begin{itemize}
    \uncover<1->{
  \item Without concurrency control, distributed edges can become reciprocally inconsistent
    \begin{itemize}
    \item A distributed edge's reciprocal information is separated by the network
    \end{itemize}
  }
  \uncover<2->{
  \item Reciprocal inconsistency is a source of corruption\footnotemark
  }
  \uncover<3->{
  \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
    }
    \end{itemize}
    \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
\end{frame}

\begin{frame}
  \frametitle{Our Solution: Collision Detection}
  \begin{itemize}
    \uncover<1->{
    \item Carefully abort transaction(s)
    }
    \uncover<2->{
    \item No centralized control or synchronized clock
    }
    \uncover<3->{
    \item Permits interleavings that preserve reciprocal consistency
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Collision Detection General Rules}
  \begin{itemize}
    \uncover<1->{
    \item Updates are initially provisional
    }
    \uncover<2->{
    \item Transactions distinguish between their first and second updates using labels
    }
    \uncover<3->{
    \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
    }
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Interference-free Update}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<1->{
        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
      }
      \uncover<2->{
        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
      }
      \uncover<3->{
        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
      }
      \uncover<3->{
        \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
      }
      \uncover<4->{
        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
      }
    \end{tikzpicture}
  \end{figure}

\end{frame}

\begin{frame}
  \frametitle{Collision Detection}
  \begin{center}
    \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
  \end{center}
  \begin{figure}[h!]
    \centering
    \begin{tikzpicture}
      \uncover<2->{
        \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
        \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
        \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
        \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
        \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
      }
      \uncover<2->{
        \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
      }
      \uncover<3->{
        \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
      }
      \uncover<4-11>{
        \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
      }
      \uncover<5-10>{
        \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
      }
      \uncover<6-10>{
        \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
      }
      \uncover<7-7>{
        \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
      }
      \uncover<8-10>{
        \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
        \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
      }
      \uncover<9-9>{
        \node [] at (6,2) {$T_R$'s 2 is missing};
      }
      \uncover<10-10>{
        \node[] at (6,2) {$T_B$ ABORTS};
      }
      \uncover<12->{
        \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
      }
      \uncover<13->{
        \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
      }
      \uncover<14->{
        \node[] at (6,2) {$T_R$ SUCCEEDS};
        }
    \end{tikzpicture}
  \end{figure}
\end{frame}

%------------------------------------------------
%Problem 2: Edge-Order Consistency
% ------------------------------------------------

\begin{frame}
  \frametitle{Edge-Order Consistency}
  A second consistency problem related to distributed edges:
  \begin{itemize}
    \uncover<2->{
    \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
    }
    \uncover<3->{
    \item Each update is reciprocally consistent
    }
    \uncover<4->{
    \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
    }
  \end{itemize}
\end{frame}

>>>>

@@

\input{config}

\input{title}

@@

\input{title-images}

@@

\input{s1-property-graph}

\input{s2-storage-layer}

\input{s3-reciprocal-consistency}

\input{s4-dist-graph}

\input{s6-half-corruption}

\input{s7-delta}

\input{s8-evaluation}

\input{s9-summary}

% %------------------------------------------------
% % Problem 1: Reciprocal Consistency}
% % ------------------------------------------------

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   Two concurrent transactions:
%   \begin{itemize}
%   \item Mr Red requests to book the seat, $T_R$
%   \item Mr Blue requests to book the seat, $T_B$
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}[node distance=2cm]
%       \uncover<1-> {
%         \draw [black,<-,>=stealth] (0,0) -- (0,4) node [] {};
%         \draw [black,<-,>=stealth] (4,0) -- (4,4) node [] {};
%         \draw [black,dashed,->,>=stealth] (0,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (4,4.1) node [label=above:{$S_2$}] {};
%       }
%       \uncover<2-3> {
%         \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (2,2.3);
%       }
%       \uncover<3-4> {
%         \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (2,2.1);
%       }
%       \uncover<4-> {
%         \draw [red,->,>=stealth] (0,3.6) node [label=left:{$T_R$}] {} -- (4,1);
%       }
%       \uncover<5-> {
%         \draw [blue,->,>=stealth] (4,3) node [label=right:{$T_B$}] {} -- (0,1.2);
%       }
%     \end{tikzpicture}
%     \caption{Concurrent transactions interleave and violate reciprocal consistency}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Reciprocal Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}[node distance=7cm]
%       \node (n1) [bignode] {\texttt{Flight}};
%       \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
%       \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=7cm,label=above:{$S_2$}] {};
%       \node (n2) [bignode, right of=n1] {\texttt{Seat}};
%       \draw [local] (n1) to[out=0,in=-180]  node[below] {\texttt{:BOOKED}} (n2);

%       \uncover<2-> {
%       \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2.5cm] {\textit{Mr Blue Booked In Seat}};
%       \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=7cm,yshift=-2.5cm] {\textit{Mr Red Booked In Seat}};
%       \draw [<->,dashed](q1) -- (q2);
%       \draw [<->,dashed](q1) -- (q2);
%       \coordinate (a) at (4,-2);
%       \coordinate (b) at (3,-3);
%       \coordinate (c) at (3,-2);
%       \coordinate (d) at (4,-3);
%       \draw [very thick,red] ($ (a) $) -- ($ (b) $);
%       \draw [very thick,red] ($ (c) $) -- ($ (d) $);
%     }
%     \end{tikzpicture}
%     \caption{Reciprocal inconsistent distributed edge}
%   \end{figure}
% \end{frame}

% % \begin{frame}
% %   \frametitle{Reciprocal Consistency Violation}
% %   \begin{figure}[h!]
% %     \centering
% %     \begin{tikzpicture}[node distance=2cm]
% %       \uncover<1->{
% %         \draw [black,<-,>=stealth] (4.5,0) -- (4.5,4) node [] {};
% %         \draw [red,->,>=stealth] (4.5,2.2) node [label=left:{$T_x$}] {} -- (6.5,1);
% %         \draw [blue,->,>=stealth] (6.5,2.7) node [label=right:{$T_y$}] {} -- (4.5,1.2);
% %         \draw [black,<-,>=stealth] (6.5,0) -- (6.5,4) node [] {};
% %         \draw [black,dashed,->,>=stealth] (4.5,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (6.5,4.1) node [label=above:{$S_2$}] {};
% %       }
% %       \uncover<2->{
% %         \draw [black,<-,>=stealth] (9,0) -- (9,4) node [] {};
% %         \draw [red,->,>=stealth] (9,3.5) node [label=left:{$T_x$}] {} -- (11,1);
% %         \draw [blue,->,>=stealth] (9,2.8) node [label=left:{$T_y$}] {} -- (11,2);
% %         \draw [black,<-,>=stealth] (11,0) -- (11,4) node [] {};
% %         \draw [black,dashed,->,>=stealth] (9,4.1) node [label=above:{$S_1$}] {} -- node [midway,above] {$E$} (11,4.1) node [label=above:{$S_2$}] {};
% %       }
% %     \end{tikzpicture}
% %     \caption{Possible interleaving of concurrent transactions that violate reciprocal consistency}
% %   \end{figure}
% % \end{frame}

% \begin{frame}
%   \frametitle{Distributed Edge Reciprocal Consistency}
%   \begin{itemize}
%     \uncover<1->{
%   \item Without concurrency control, distributed edges can become reciprocally inconsistent
%     \begin{itemize}
%     \item A distributed edge's reciprocal information is separated by the network
%     \end{itemize}
%   }
%   \uncover<2->{
%   \item Reciprocal inconsistency is a source of corruption\footnotemark
%   }
%   \uncover<3->{
%   \item No known concurrency control protocol exists specific to graph databases and maintaining reciprocal consistency
%     }
%     \end{itemize}
%     \footnotetext[3]{Ezhilchelvan, P. et al, \textit{On the degradation of distributed graph databases with eventual consistency}. European Performance Engineering Workshop 2018.}
% \end{frame}

% \begin{frame}
%   \frametitle{Our Solution: Collision Detection}
%   \begin{itemize}
%     \uncover<1->{
%     \item Carefully abort transaction(s)
%     }
%     \uncover<2->{
%     \item No centralized control or synchronized clock
%     }
%     \uncover<3->{
%     \item Permits interleavings that preserve reciprocal consistency
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Collision Detection General Rules}
%   \begin{itemize}
%     \uncover<1->{
%     \item Updates are initially provisional
%     }
%     \uncover<2->{
%     \item Transactions distinguish between their first and second updates using labels
%     }
%     \uncover<3->{
%     \item When a transaction attempts to update a given record, it can identify all other transactions that have earlier updated that record provisionally
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Interference-free Update}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<1->{
%         \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
%         \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
%         \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
%       }
%       \uncover<2->{
%         \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
%       }
%       \uncover<3->{
%         \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
%       }
%       \uncover<3->{
%         \draw [red,->,>=stealth] (0,3.5) node [] {} -- (4,0.3) node (TX) [] {};
%       }
%       \uncover<4->{
%         \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
%       }
%     \end{tikzpicture}
%   \end{figure}

% \end{frame}

% \begin{frame}
%   \frametitle{Collision Detection}
%   \begin{center}
%     \textbf{Rule: For any ``1'' seen, there must be a ``2'' in the opposite end. Else, abort}
%   \end{center}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<2->{
%         \draw [black,<-,>=stealth] (-3,1) -- (-3,3) node [label=above:{time}] {};
%         \draw [black,<-,>=stealth] (0,0) -- (0,4) ;
%         \draw [black,<-,>=stealth] (4,0) -- (4,4) ;
%         \draw [black,dashed,->,>=stealth] (0,4.5) node [label=above:{$e_{out}$ in $S_1$}] {} -- (4,4.5) node [label=above:{$e_{in}$ in $S_2$}] {};
%         \draw [white,<-,>=stealth] (7,1) -- (7,3) ; %placeholder to balance picture
%       }
%       \uncover<2->{
%         \node [] at (-0.5,4) {\textcolor{red}{$T_R$}};
%       }
%       \uncover<3->{
%         \node [] at (-0.5,3.5) {\textcolor{red}{$1$}};
%       }
%       \uncover<4-11>{
%         \draw [red,->,>=stealth] (0,3.5) -- (2,1.9);
%       }
%       \uncover<5-10>{
%         \node [] at (-0.5,2.6) {\textcolor{blue}{$T_B$}};
%       }
%       \uncover<6-10>{
%         \node [] at (-0.5,2.1) {\textcolor{blue}{$1$}};
%       }
%       \uncover<7-7>{
%         \draw [blue,->,>=stealth] (0,2.1) -- (2,1.8);
%       }
%       \uncover<8-10>{
%         \draw [blue,->,>=stealth] (0,2.1) -- (4,1.5);
%         \node [] at (4.5,1.5) {\textcolor{blue}{$2$}};
%       }
%       \uncover<9-9>{
%         \node [] at (6,2) {$T_R$'s 2 is missing};
%       }
%       \uncover<10-10>{
%         \node[] at (6,2) {$T_B$ ABORTS};
%       }
%       \uncover<12->{
%         \draw [red,->,>=stealth] (0,3.5)  -- (4,0.3);
%       }
%       \uncover<13->{
%         \node [] at (4.5,0.3) {\textcolor{red}{$2$}};
%       }
%       \uncover<14->{
%         \node[] at (6,2) {$T_R$ SUCCEEDS};
%         }
%     \end{tikzpicture}
%   \end{figure}
% \end{frame}

% %------------------------------------------------
% %Problem 2: Edge-Order Consistency
% % ------------------------------------------------

% \begin{frame}
%   \frametitle{Edge-Order Consistency}
%   A second consistency problem related to distributed edges:
%   \begin{itemize}
%     \uncover<2->{
%     \item Consider 2 transactions that update $2$ or more of the \textit{same} distributed edges
%     }
%     \uncover<3->{
%     \item Each update is reciprocally consistent
%     }
%     \uncover<4->{
%     \item Then if $T_R$ updates before $T_B$ on a given edge then this order should be preserved across all edges
%     }
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge-Order Consistency Violation}
%   Consider two edges,
%   \begin{figure}[h!]
%     \centering
%     \resizebox{1\textwidth}{!}{
%       \begin{tikzpicture}[node distance=5.2cm]
%         \node[] at (2.6,2.5) {\large{$E$}};
%         \node[] at (13,2.5) {\large{$E'$}};
%         \node (n1) [bignode] {\texttt{Flight}};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label=above:{$S_1$}] {};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=5.2cm,label=above:{$S_2$}] {};
%         \node (n2) [bignode, right of=n1] {\texttt{Seat}};
%         \draw [local] (n1) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n2);
%         \node (q1) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=0cm,yshift=-2cm] {\textit{Seat Available In Flight}};
%         \node (q2) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=5.2cm,yshift=-2cm] {\textit{In Flight Seat Available}};
%          \draw [<->,dashed](q1) -- (q2);

%         \node (n3) [bignode, right of=n2] {\texttt{Hotel}};
%         \node (n4) [bignode, right of=n3] {\texttt{Room}};
%         \draw [local] (n3) to[out=0,in=-180]  node[below] {\small{\texttt{:AVAILABLE}}} (n4);
%         \node (q3) [rectangle, minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=10.4cm,yshift=-2cm] {\textit{Room Available In Hotel}};
%         \node (q4) [rectangle,minimum width=0.6cm,minimum height=0.6cm,draw=black,xshift=15.6cm,yshift=-2cm] {\textit{In Hotel Room Available}};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,xshift=15.6cm,label=above:{$S_4$}] {};
%         \node (rect)  [draw,dashed,minimum width=3cm,minimum height=3cm,label,xshift=10.4cm,label=above:{$S_3$}] {};
%         \draw [<->,dashed](q3) -- (q4);

%       \end{tikzpicture}
%     }
%     \caption{Two distributed edges}
%   \end{figure}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge-Order Consistency Violation}
%   Two concurrent transactions by a travel agent:
%   \begin{itemize}
%   \item Requests to book room and seat for Mr Red, $T_R$
%   \item Requests to book room and seat for Mr Blue, $T_B$
%   \end{itemize}
% \end{frame}

% \begin{frame}
%   \frametitle{Edge-Order Consistency Violation}
%   \begin{figure}[h!]
%     \centering
%     \begin{tikzpicture}
%       \uncover<1->{

%         \draw [black,->,>=stealth] (0,0) -- (10,0) node [label=below:{$time$}] {};
%         \draw [black,-,>=stealth] (2,4) node [label={[red]left:{$T_R$}}] {} -- (8,4);
%         \draw [black,-,>=stealth] (8,4.1) node [] {} -- (8,3.9);
%         \draw [black,-,>=stealth] (2,4.1) node [] {} -- (2,3.9);
%         \draw [black,-,>=stealth] (4,2) node [label={[blue]left:{$T_B$}}] {} -- (7,2);
%         \draw [black,-,>=stealth] (4,2.1) node [] {} -- (4,1.9);
%         \draw [black,-,>=stealth] (7,2.1) node [] {} -- (7,1.9);
%       }
%       \uncover<2->{
%         \draw [black,dashed] (2.5,0) node [label=below:{$t_1$}] {} -- (2.5,4) node [label=above:{$E$}] {};
%       }
%       \uncover<3->{
%         \draw [black,dashed] (4.5,0) node [label=below:{$t_2$}] {} -- (4.5,2) node [label=above:{$E$}] {};
%       }
%       \uncover<4->{

%         \draw [black,dashed] (6.5,0) node [label=below:{$t_3$}] {} -- (6.5,2) node [label=above:{$E'$}] {};
%       }
%       \uncover<5->{

%         \draw [black,dashed] (7.5,0) node [label=below:{$t_4$}] {} -- (7.5,4) node [label=above:{$E'$}] {};
%       }
%       \uncover<6->{
%         \draw (8.3,2) node [label=right:{$E : T_R \rightarrow T_B$}] {} ;
%         \draw (8.3,1.5) node [label=right:{$E' : T_B \rightarrow T_R$}] {} ;
%         \draw (8.3,1) rectangle (11,2.5);
%       }
%     \end{tikzpicture}
%     \caption{Edge-order consistency violation}
%   \end{figure}
% \end{frame}
